\mytitle{}

### 2018-07-13 paper/paper.tex
Far more fleshed out. Gaps very rel.

\usepackage[binary-units, per-mode=symbol]{siunitx}
\sisetup{range-phrase=--, range-units=single}

@@

\usepackage{booktabs}
\usepackage{microtype}

@@

Network anomaly detection, intrusion detection and intrusion prevention are continually evolving problems, compounded by the partial, non-IID view of data at each point in the network.
Attacks and anomalous behaviours evolve, becoming more sophisticated or employing new vectors to harm a network or system's confidentiality, integrity, or availability without being detected.
These attacks and anomalies have measurable consequences and symptoms which allow a skilled analyst to infer new signatures for detection by misuse-based classifiers, but unseen attacks may only be defended against after-the-fact.

While \emph{machine learning} (ML) approaches seem like a sensible fit for this problem, in \citeyear{DBLP:conf/sp/SommerP10} \textcite{DBLP:conf/sp/SommerP10} identified the `failure to launch' of ML-based anomaly detection systems---to quite a large extent, this remains the case today.
Their application is made difficult due to significant operational differences from standard ML tasks, alongside certain characteristics of network traffic.
Briefly, these are the diversity of network traffic across varying timescales \cite{DBLP:conf/sp/SommerP10} and significant burstiness \cite{DBLP:journals/ccr/LelandWTW95}.
Above the aggregate level, the constant deployment of new services and new protocols means that traffic is \emph{non-stationary} and displays an evolving notion of normality.
This is compounded by the extraordinarily low tolerance for false positives inherent to network intrusion detection \cite{DBLP:conf/ccs/Axelsson99}, and the challenges of learning from unlabelled (often partial) data.
All of these factors greatly inflate the difficulty of the detection problem.

For certain classes of problem like non-transit DDoS attacks, \emph{reinforcement learning} (RL) is best-posed to adaptively react to certain classes of threat by assuming the role of a feedback loop for network optimisation, typically to safeguard service guarantees.
In a sense, this allows us to ``overcome'' the difficulties of the detection problem by monitoring \emph{performance characteristics and consequences} in real-time, by looking for the effect rather than the cause.
The intent is to augment what existing solutions can provide, by automatically alerting, recording and controlling what are believed to be illegal system states.
%Whether it takes direct control of the network, or is used indirectly to optimise a key part of another system, more powerful `deep' RL techniques (and well-founded action spaces) aren't yet well explored for network IDS/IPS.
%These range from more modern training algorithms \cite{DBLP:journals/corr/SchulmanWDRK17, DBLP:conf/icml/SchulmanLAJM15}, to evolutionary strategies \cite{DBLP:journals/corr/SalimansHCS17, DBLP:journals/corr/abs-1802-08842}, hierarchical action composition \cite{DBLP:journals/corr/abs-1710-09767}, and competitive multi-agent learning \cite{DBLP:journals/corr/abs-1710-03748}.

@@

\section{Background}

%?? Introduce RL, related definitions etc.
\emph{Reinforcement learning} (RL) is a variant of machine learning principally concerned with with training an agent to choose an optimal sequence of actions in the pursuit of a given task \cite{RL2E}.
We assume the agent has a certain amount of knowledge whenever a decision must be made: at any point in time $t$ it knows which \emph{state} it is in ($S_t \in \mathcal{S}$), the set of \emph{actions} which are available to it ($\operatorname{A}(S_t) \subseteq \mathcal{A}$) and, if available, a numeric \emph{reward} obtained from the last action chosen ($R_t \in \mathbb{R}, A_{t-1} \in \operatorname{A}(S_{t-1})$).
This formulation of interaction with a system is best described as a \emph{Markov decision problem} (MDP).
RL methods combine this information with a current \emph{policy} $\pi$ to determine which action should be taken: such a choice need not be optimal, if an agent needs to further explore some region of the state space.
The policy is then further refined by updating value estimates for state-action pairs or via policy gradient methods, meaning that RL-based approaches learn adaptively and online if reward functions are available in their deployment environment.
From any point in a sequence of decisions, we may describe the sum of rewards yet to come as the \emph{discounted return}, $G_t = R_{t+1} + \gamma R_{t+2} + \gamma^2 R_{t+2} + \ldots$, choosing the discount factor $\gamma \in [0,1]$ to determine how crucial future rewards are vis-\`{a}-vis the current state.
Formally then, the goal is to choose actions which maximise the \emph{expected discounted return} $\operatorname{\mathbb{E}_{\pi}}[G_t]$.

@@

There is immense variation in \emph{how} policies and/or values may be learned, reliant upon the learning environment, problem and required convergence guarantees.
In particular, we shall focus on methods which choose actions according to their value estimates from the current state.
To cope with a continuous state- and/or action-space, one valuable technique is to employ \emph{tile-coding} \cite[pp.\ \numrange{217}{221}]{RL2E}---this converts our state from $\mathbb{R}^d \times \mathcal{A}$ into a sparse boolean feature vector $\operatorname{\mathbf{x}}(s, a)$ by subdividing it into a number of overlapping $(d+|\mathcal{A}|)$-dimensional grids with an optional bias component.
Entries of $\operatorname{\mathbf{x}}(s, a)$ are set to 1 if the state action pair exists within the corresponding tile.
This transform, as an example of linear function approximation, then necessitates the use of \emph{semi-gradient Sarsa} \cite[pp.\ \numrange{243}{244}]{RL2E}.
Given a weight vector $\wvec{0}=\bm{0}$ and learning rate $\alpha$, we may approximate an action $a$'s value in any state $s$ by $\operatorname{q}(s, a) \approx \acval{s}{a}{\wvec{}}$:
\begin{subequations}
	\begin{gather}
	\acval{s}{a}{\wvec{}} = \wvec{}^{\top} \operatorname{\mathbf{x}}(s, a),\\
	\delta_t = R_{t+1} + \gamma \acval{S_{t+1}}{A_{t+1}}{\wvec{t}} - \acval{S_t}{A_t}{\wvec{t}},\\
	\bm{w}_{t+1} = \bm{w}_{t} + \alpha \delta_t \nabla{\acval{S_t}{A_t}{\wvec{t}}}.
	\end{gather}
	\label{eqn:sg-sarsa}
\end{subequations}
By computing the approximate value of every action available in the current state, actions with maximal value can be chosen each time (the \emph{greedy} policy), we might take a random action with probability $\epsilon$ to encourage early exploration (the \emph{$\epsilon$-greedy policy}), or we might some use some other mechanism.

If $\operatorname{A}(s) = \mathcal{A}$ in every state (and actions are themselves discrete), then the implementation of the policy and update rules may be simplified by storing a vector of action values for each state encountered.
Observing that $\nabla{\acval{s}{a}{\wvec{}}} = \operatorname{\mathbf{x}}(s, a)$, we may gain further optimisations by considering that a tile-coded feature vector is a boolean vector of constant Hamming weight (and so is amenable to representation as an array of indices).
Moreover, we need not store vectors of action values for tiles which have not yet been visited.

@@

\section{A Plan, of Sorts}

\begin{enumerate}
	\item The main case for contribution in what I have so far:
	\begin{itemize}
		\item Past work reliant on unrealistic network models: tcp-like behaviour (and its effects on collateral damage) not captured, disjoint ranges of traffic distribution (no benign heavy-hitters), ISP-like topology.
		\item I offer more realistic network emulation environment, better treatment of protocol/traffic characteristics.
	\end{itemize}
	\item Forthcoming: rethinking state/action spaces to operate at a finer level of granularity. New network model (live tcp back-and-forth), allows us to test collateral damage assumptions in a more realistic manner (and show clear case for moving beyond work of malialis and kudenko)
\end{enumerate}

@@

%\draw[-] (teaml) -- (endteaml);
\node [labeltexts] (kdots) at ($(teaml)!0.5!(endteaml)$) {$\rvdots$};
\node [labeltexts, right = -0.1cm of kdots] {$k$};
\node [labeltexts] (ldots) at ($(inter)!0.5!(endinter)$) {$\rvdots$};
\node [labeltexts, right = -0.1cm of ldots] {$l$};
\node [labeltexts] (mdots) at ($(agent)!0.5!(endagent)$) {$\rvdots$};
\node [labeltexts, right = -0.1cm of mdots] {$m$};
\node [labeltexts] (ndots) at ($(host)!0.5!(endhost)$) {$\rvdots$};
\node [labeltexts, right = -0.1cm of ndots] {$n$};
\end{tikzpicture}
}
\caption{
	Network topology diagram, showing how the server and its core switch's $k$ teams are structured, with $l$ intermediate routers per team, connected to $m$ agents which each moderate $n$ hosts beyond a single external switch.
	Unfilled nodes are considered to be internal.
	\label{fig:marl-topol}
}
\end{figure}

@@

At every time step, each agent installs an action via OpenFlow, instructing its host switch to drop each packet with probability $p$.
Specifically, the agent chooses $p \in \left\{ 0.0, 0.1, \ldots, 0.9 \right\}$, giving a discrete, static action set which cannot completely filter traffic.

@@

The tile-coded state and chosen reward function are then used as input for the \emph{semi-gradient Sarsa} algorithm described in \cref{eqn:sg-sarsa}, making use of $\epsilon$-greedy action selection.

@@

\section{Performance in an Emulated Environment}

The original introduction of this approach to direct-control reinforcement learning as introduced by \textcite{DBLP:journals/eaai/MalialisK15} fails to consider key cases: the accurate simulation of TCP-like behaviour (and its effects on collateral damage), disjoint ranges of traffic distribution (i.e., the presence of benign heavy-hitters), and assumes a rigid ISP-like topology.

@@

\subsection{Service Guarantees and Host Density}
The .
In practice, well-established network topologies such as the \emph{fat-tree topology} fan out internally (to provide efficient replication and scaling of resources) while having comparatively few egress points.
It is this latter condition which I feel is most worth examining: the case where the host/agent ratio is heavily skewed in favour of hosts, and actions as they have been defined are forced to become less granular.

The base `online' experiment I replicate configures the network topology using $k=5$ teams, $l=3$ intermediate nodes per team, $m=2$ agents per intermediate node, and $n \in \{2, 7, 11, 14\}$ hosts per learner.
I set $\gamma=0$ (leading to opportunistic behaviour), $\alpha=0.05$ and linearly annealing $\epsilon=0.2 \rightarrow 0$ by $t=3000$.
Benign and malicious hosts uploaded between \SIrange{0}{1}{\mega\bit\per\second} and \SIrange{2.5}{6}{\mega\bit\per\second} respectively, and hosts are redrawn at each episode's start with $\operatorname{P}(\mathit{malicious})=0.4$.
$U_s$ was fixed at $klmn+2\si{\mega\bit\per\second}$.
The performance of each choice of $n$ was averaged over \num{10} episodes of length \num{10000} timesteps (setting $\wvec{}=\bm{0}$ between episodes).
Host allocations were generated pseudorandomly to ensure fairness between choices of $n$.

\begin{figure}
	\includegraphics[width=\linewidth]{../plots/online-varyN-uneven}
	\caption{
		MARL pushback control system performance plotted over time, for various settings of $n$ hosts per learner.
		This plot shows that, from the perspective of benign hosts, service guarantees degrade as inference and applied actions become less granular.
		This generalises for all behavioural discriminators: even a perfect agent \emph{must} punish benign flows if they are grouped with malicious actors.
		%		?? Effect on max perf AND learning rate??
		Crucially, both the maximum achievable performance and learning rate are negatively impacted when granularity is coarse.
		\label{fig:marl-granularity}
	}
\end{figure}

\Cref{fig:marl-granularity} shows that allocating more hosts to each learner in the recreated experiment leads to a decrease in the amount of benign traffic observed at the server.
Given the UDP-like behaviour of Tcpreplay, being able to replicate the online performance of \citeauthor{DBLP:journals/eaai/MalialisK15} in the $n=2$ case suggests that the numeric simulation their work is entirely based upon does not capture the key interactions between TCP flows and packet drop as a control action.

Notably, modelling the environment in Mininet makes evaluation rather slow as any interaction, coordination and communication must take place in real-time.
While I'm sure this enhances the quality of model learned due to the explicit inclusion of more accurate network behaviour, a more abstract simulator based on observed physical properties could provide data at less computational cost.

\subsection{TCP Multiplexing and Fallback Behaviour}
While Tcpreplay was useful in recreating the original work, flow behaviour under packet loss is closer to that of UDP when using it to replay traffic.
To validate the operation of the pdrop OpenFlow action and to investigate the effects of collateral damage upon benign TCP flows, I conducted an iperf3 test using mininet.
Several hosts communicated concurrently to a single server over one switch, with maximum send rates in $\{1,...,5\}$ \si{\mega\bit\per\second} and a constant \SI{40}{\milli\second} latency.
During the test, the rate of packet drop was increased linearly from \SIrange{0}{40}{\percent} and the observed throughput for each host was measured.

@@

\Cref{fig:mplex-tcp} shows that benign TCP flows are severely harmed by \emph{any} packet drop, so collateral damage to these hosts is significant.
Good-faith TCP congestion avoidance causes these flows to attempt to back off very quickly.
%?? Talk about the same major points you made in the presentation here (i.e. TCP interactions) w/ pushback. Or in the evaluation...
UDP, however, behaves as expected---malicious hosts are expected to exhibit the same behaviour, given that congestion is their primary goal.

?? TCP perf characteristics under packet drop known by the Mathis equation \cite{DBLP:journals/ccr/MathisSMO97}:
\begin{equation}
\operatorname{BW} = \frac{\operatorname{MSS}}{\operatorname{RTT}} \frac{C}{\sqrt{p}},
\end{equation}
for the \emph{maximum segment size} (MSS), \emph{round trip time} (RTT), a choice of constant $C \approx{} \sqrt{3/2}$ dependent on system assumptions and the probability $p \in (0, 1]$ that a packet is lost.
The lower bound on $p$ is deliberately open, as packet drop acts as a control signal for the TCP send window size and is an expected part of normal operation.

\subsection{Synthesis}

%It'd be really good if I could figure out a rough proof for the service property...
%Okay... now say what this means for future viability! What changes can I make? Flow-specific metrics, of course! Increase granularity as much as I can!
Both of these experiments suggest that the most viable way to enhance and deploy this MARL approach is to consider flows individually to maximise performance for legitimate hosts and to detect malicious hosts in a more principled manner.
While this will add significant computational cost, this may be worked around with intelligent sampling and monitoring while taking multiple actions per-timestep \cite{DBLP:conf/hotnets/MaoAMK16}.

?? Observations: training time lengthier by nature of emulated environment.
?? Risk of going for a simulation (i.e. numeric)? Always going to be interesting behaviour that is missed out on (in theory), at the cost of training time/limits of simulation speed

@@

?? Other recent work in the prevention of DDoS? (i.e., non-RL).

@@

?? Good Direct-control work \cite{DBLP:phd/ethos/Malialis14, DBLP:journals/eaai/MalialisK15}

?? Indirect-control applications: link to all the juicy stuff in e.g. HotNets, data-driven networking, the works. \cite{DBLP:conf/hotnets/ValadarskySST17, DBLP:conf/hotnets/MaoAMK16}.
The HMM stuff---learning when best to \emph{communicate} and share knowledge between explicit detector models \cite{DBLP:conf/paisi/XuSH07}.
The latter example's position is slightly weakened by its reliance on the discredited `DARPA99' dataset \cite{DARPA-IDD, DBLP:conf/cisda/TavallaeeBLG09, DBLP:conf/sp/SommerP10}, but the idea itself is well-treated and this acts as a driver for improvements in this direction.

@@

It all ends here, Mr.\ Bond...

@@

?? Benefit of the more realistic emulation environ is that it is far closer in behaviour and architecture (i.e. viable) to a real SDN-enabled deployment, captures some dynamics which were otherwise hidden/lost by human ignorance. It also allows me to develop the system towards evolving traffic models where it is expected that RL should shine over and above standard ML techniques. Room to introduce/roll-in dynamic changepoint detection or adaptive exploration \cite{DBLP:conf/ki/Tokic10, DBLP:conf/ki/TokicP11, DBLP:conf/annpr/TokicP12}?

?? Overarching goal: works which respect the complexity of the network environment (rather than na\"{i}ve simulation, blind ML applications etc.) and choose well-considered pathways to solution. Call-to-action?

?? Security? I suspect that the very qualities that make inference difficult in IDS/IPS also increase the level of challenge an advanced threat must overcome.
System state which is dependent on many signals drawn from across a wide network, in concert with their burstiness and unpredictability, may have substantial effects on an attacker's capabilities.

>>>>

\usepackage[binary-units, per-mode=symbol]{siunitx}
\sisetup{detect-all, range-phrase=--, range-units=single}

@@

\usepackage{booktabs}
\usepackage[
activate={true,nocompatibility},
final,
tracking=true,
%kerning=true,spacing=true
]{microtype}
\microtypecontext{spacing=nonfrench}

@@

Network anomaly detection, intrusion detection and intrusion prevention are continually evolving problems, compounded by the partial, non-IID view of data at each point in the network.
Attacks and anomalous behaviours evolve, becoming more sophisticated or employing new vectors to harm a network or system's confidentiality, integrity, or availability without being detected.
These attacks and anomalies have measurable consequences and symptoms which allow a skilled analyst to infer new signatures for detection by misuse-based classifiers, but unseen attacks may only be defended against after-the-fact.
This issue is inherent to \emph{misuse-based} intrusion detectors, and it has been long-hoped that \emph{anomaly-based} detectors would surpass this by making use of statistical measures.

While \emph{machine learning} (ML) approaches seem like a sensible fit for this problem, in \citeyear{DBLP:conf/sp/SommerP10} \textcite{DBLP:conf/sp/SommerP10} identified the `failure to launch' of ML-based anomaly detection systems---to quite a large extent, this remains the case today.
Their application is made difficult due to significant operational differences from standard ML tasks, alongside certain characteristics of network traffic.
Briefly, these are the diversity of network traffic across varying timescales \cite{DBLP:conf/sp/SommerP10} and significant burstiness \cite{DBLP:journals/ccr/LelandWTW95}.
Above the aggregate level, the constant deployment of new services and new protocols means that traffic is \emph{non-stationary} and displays an evolving notion of normality.
Further issues arise from the extraordinarily low tolerance for false positives inherent to network intrusion detection \cite{DBLP:conf/ccs/Axelsson99}, and the challenges of learning from unlabelled (often partial) data.
All of these factors greatly inflate the difficulty of the detection problem.

For certain classes of problem like flooding-based DDoS attacks, \emph{reinforcement learning} (RL) offers another perspective for adaptively reacting to certain classes of threat by assuming the role of a feedback loop for network optimisation, typically to safeguard service guarantees.
In a sense, this allows us to ``overcome'' the difficulties of the detection problem by monitoring \emph{performance characteristics and consequences} in real-time, by looking for the effect rather than the cause.
The intent is to augment what existing misuse-based solutions can provide, by automatically alerting, recording and controlling what are believed to be illegal system states.
%Whether it takes direct control of the network, or is used indirectly to optimise a key part of another system, more powerful `deep' RL techniques (and well-founded action spaces) aren't yet well explored for network IDS/IPS.
%These range from more modern training algorithms \cite{DBLP:journals/corr/SchulmanWDRK17, DBLP:conf/icml/SchulmanLAJM15}, to evolutionary strategies \cite{DBLP:journals/corr/SalimansHCS17, DBLP:journals/corr/abs-1802-08842}, hierarchical action composition \cite{DBLP:journals/corr/abs-1710-09767}, and competitive multi-agent learning \cite{DBLP:journals/corr/abs-1710-03748}.

@@

\section{Background}
%?? Introduce RL, related definitions etc.
\emph{Reinforcement learning} (RL) is a variant of machine learning principally concerned with with training an agent to choose an optimal sequence of actions in the pursuit of a given task \cite{RL2E}.
We assume the agent has a certain amount of knowledge whenever a decision must be made: at any point in time $t$ it knows which \emph{state} it is in ($S_t \in \mathcal{S}$), the set of \emph{actions} which are available to it ($\operatorname{A}(S_t) \subseteq \mathcal{A}$) and, if available, a numeric \emph{reward} obtained from the last action chosen ($R_t \in \mathbb{R}, A_{t-1} \in \operatorname{A}(S_{t-1})$).
This formulation of interaction with a system is best described as a \emph{Markov decision process} (MDP).
RL methods combine this information with a current \emph{policy} $\pi$ to determine which action should be taken: such a choice need not be optimal, if an agent needs to further explore some region of the state space.
The policy is then further refined by updating value estimates for state-action pairs or via policy gradient methods, meaning that RL-based approaches learn adaptively and online if reward functions are available in their deployment environment.
From any point in a sequence of decisions, we may describe the sum of rewards yet to come as the \emph{discounted return}, $G_t = R_{t+1} + \gamma R_{t+2} + \gamma^2 R_{t+2} + \ldots$, choosing the discount factor $\gamma \in [0,1]$ to determine how crucial future rewards are vis-\`{a}-vis the current state.
Formally then, the goal is to choose actions which maximise the \emph{expected discounted return} $\operatorname{\mathbb{E}_{\pi}}[G_t]$.

@@

There is immense variation in \emph{how} policies and/or values may be learned, reliant upon the learning environment, problem and required convergence guarantees.
In particular, we shall focus on methods which choose actions according to their value estimates from the current state.
To cope with a continuous state- and/or action-space, one valuable technique is to employ \emph{tile-coding} \cite[pp.\ \numrange{217}{221}]{RL2E}---this converts our state from $\mathbb{R}^d \times \mathcal{A}$ into a sparse boolean feature vector $\operatorname{\mathbf{x}}(s, a)$ by subdividing it into a number of overlapping $(d+ \dim{\mathcal{A}})$-dimensional grids with an optional bias component.
Entries of $\operatorname{\mathbf{x}}(s, a)$ are set to 1 if the state-action pair exists within the corresponding tile.
This transform, as an example of linear function approximation, then necessitates the use of \emph{1-step semi-gradient Sarsa} \cite[pp.\ \numrange{243}{244}]{RL2E}.
Given a weight vector $\wvec{0}=\bm{0}$ and learning rate $\alpha$, we may approximate an action $a$'s value in any state $s$ taking $\operatorname{q}(s, a) \approx \acval{s}{a}{\wvec{}}$ by continually updating $\wvec{t}$ as follows:
\begin{subequations}
	\begin{gather}
	\acval{s}{a}{\wvec{}} = \wvec{}^{\top} \operatorname{\mathbf{x}}(s, a),\\
	\delta_t = R_{t+1} + \gamma \acval{S_{t+1}}{A_{t+1}}{\wvec{t}} - \acval{S_t}{A_t}{\wvec{t}},\\
	\bm{w}_{t+1} = \bm{w}_{t} + \alpha \delta_t \nabla{\acval{S_t}{A_t}{\wvec{t}}},
	\end{gather}
	\label{eqn:sg-sarsa}
	taking $\nabla$ with respect to $\wvec{}$.
\end{subequations}

Computing the approximate value of every action available in the current state forms the basis of a policy.
Actions with maximal value can be chosen each time (the \emph{greedy} policy), we might modify this by taking random actions with probability $\epsilon$ to encourage early exploration (the \emph{$\epsilon$-greedy policy}), or we might some use some other mechanism.

If $\operatorname{A}(s) = \mathcal{A}$ in every state (and actions are themselves discrete), then the implementation of the policy and update rules may be simplified by storing a vector of action values for each state encountered.
Observing that $\nabla{\acval{s}{a}{\wvec{}}} = \operatorname{\mathbf{x}}(s, a)$, we may gain further optimisations by considering that a tile-coded feature vector is a binary vector of constant Hamming weight (and so is amenable to representation as an array of indices).
Action values are then simply the sum of all activated tile weight vectors.
Moreover, we need not store action values for tiles which have not yet been visited.

@@

%\section{A Plan, of Sorts}
%
%\begin{enumerate}
%	\item The main case for contribution in what I have so far:
%	\begin{itemize}
%		\item Past work reliant on unrealistic network models: tcp-like behaviour (and its effects on collateral damage) not captured, disjoint ranges of traffic distribution (no benign heavy-hitters), ISP-like topology.
%		\item I offer more realistic network emulation environment, better treatment of protocol/traffic characteristics.
%	\end{itemize}
%	\item Forthcoming: rethinking state/action spaces to operate at a finer level of granularity. New network model (live tcp back-and-forth), allows us to test collateral damage assumptions in a more realistic manner (and show clear case for moving beyond work of malialis and kudenko)
%\end{enumerate}

@@

%\draw[-] (teaml) -- (endteaml);
\node [labeltexts] (kdots) at ($(teaml)!0.5!(endteaml)$) {$\rvdots$};
\node [labeltexts, right = -0.1cm of kdots] {$k$};
\node [labeltexts] (ldots) at ($(inter)!0.5!(endinter)$) {$\rvdots$};
\node [labeltexts, right = -0.1cm of ldots] {$l$};
\node [labeltexts] (mdots) at ($(agent)!0.5!(endagent)$) {$\rvdots$};
\node [labeltexts, right = -0.1cm of mdots] {$m$};
\node [labeltexts] (ndots) at ($(host)!0.5!(endhost)$) {$\rvdots$};
\node [labeltexts, right = -0.1cm of ndots] {$n$};
\end{tikzpicture}
}
\caption{
	Network topology diagram, showing how the server and its core switch's $k$ teams are structured, with $l$ intermediate routers per team, connected to $m$ agents which each moderate $n$ hosts beyond a single external switch.
	Empty nodes are considered to be internal.
	\label{fig:marl-topol}
}
\end{figure}

@@

At every time step, each agent installs an action via OpenFlow, instructing its host switch to drop each packet with probability $p$.
Specifically, the agent chooses $p \in \left\{ 0.0, 0.1, \ldots, 0.9 \right\}$, giving a discrete, static action set which cannot completely filter traffic.
In essence, this means that each agent is in control of pushback \cite{DBLP:journals/ccr/MahajanBFIPS02a}.

@@

The tile-coded state and chosen reward function are then used as input for the \emph{semi-gradient Sarsa} algorithm described in \cref{eqn:sg-sarsa}, making use of $\epsilon$-greedy action selection.
Each agent has its own internal parameter vector $\wvec{}$, and agents do not share their experience or weight vector updates with one another.

@@

\section{Performance in an Emulated Environment}
The original introduction of this approach to direct-control reinforcement learning as introduced by \textcite{DBLP:journals/eaai/MalialisK15} fails to consider key cases: disjoint ranges of traffic distribution (i.e., the presence of benign heavy-hitters), the accurate simulation of TCP-like behaviour (and its effects on collateral damage), and assumes a rigid ISP-like topology.
The latter two are most deserving of a closer investigation, as they have stronger implications for wide-scale deployment.

@@

?? TODO: Make repo public...

\subsection{Service Guarantees and Host Density}
In the default topology, $n=2$ hosts per egress point is not a parameter that can be described as realistic for many web service deployments.
In practice, well-established network topologies such as the fat-tree topology fan out \emph{internally} (to provide efficient replication and scaling of resources) while having comparatively few egress points.
I feel is most worth examining aspects of such a case, where the host/agent ratio is heavily skewed in favour of hosts and actions as defined become less granular.

To test this, I configured the network topology using $k=2$ teams, $l=3$ intermediate nodes per team, $m=2$ agents per intermediate node, and $n \in \{2, 7, 11, 14\}$ hosts per learner.
I set $\gamma=0$ (leading to opportunistic behaviour), $\alpha=0.05$ and linearly annealed $\epsilon=0.2 \rightarrow 0$ by $t=3000$.
Benign and malicious hosts uploaded between \SIrange{0}{1}{\mega\bit\per\second} and \SIrange{2.5}{6}{\mega\bit\per\second} respectively, and hosts were redrawn at each episode's start with $\operatorname{P}(\mathit{malicious})=0.4$.
$U_s$ was fixed at $klmn+2$ \si{\mega\bit\per\second}.
The performance of each choice of $n$ was averaged over \num{10} episodes of length \num{10000} timesteps (setting each agent's $\wvec{}=\bm{0}$ between episodes).
Host allocations were generated pseudorandomly to ensure fairness between choices of $n$.

\begin{figure}
	\includegraphics[width=\linewidth]{../plots/online-varyN-uneven}
	\caption{
		MARL pushback control system performance plotted over time, for various settings of $n$ hosts per agent.
		This plot shows that, from the perspective of benign hosts, service guarantees degrade as inference and applied actions become less granular.
		This generalises for all behavioural discriminators: even a perfect agent \emph{must} punish benign flows if they are grouped with malicious actors.
		%		?? Effect on max perf AND learning rate??
		Crucially, both the maximum achievable performance and learning rate are negatively impacted when granularity is coarse.
		\label{fig:marl-granularity}
	}
\end{figure}

\Cref{fig:marl-granularity} shows that allocating more hosts to each learner leads to a guaranteed decrease in the amount of benign traffic observed at the server.
This manifests in two ways: the best-achievable performance decreases, as does the learning rate.
Given the UDP-like behaviour of Tcpreplay, being able to replicate the online performance of \citeauthor{DBLP:journals/eaai/MalialisK15} in the $n=2$ case suggests that the numeric simulation their work is entirely based upon does not capture the key interactions between TCP flows and packet drop as a control action.

\subsection{TCP Multiplexing and Fallback Behaviour}
While Tcpreplay was useful in recreating the original work and examining the above condition, flow behaviour under packet loss is closer to that of UDP when using it to replay traffic.
To investigate the effects of collateral damage upon benign TCP flows, an iperf3 test was conducted using mininet.
Several hosts communicated concurrently to a single server over one switch, with maximum send rates in $\{1,...,5\}$ \si{\mega\bit\per\second} and a constant \SI{40}{\milli\second} latency.
During the test, the rate of packet drop was increased linearly from \SIrange{0}{40}{\percent} and the observed throughput for each host was measured.

@@

\Cref{fig:mplex-tcp} shows that benign TCP flows are severely harmed by \emph{any} packet drop, so collateral damage to these hosts is more significant.
Good-faith TCP congestion avoidance causes these flows to attempt to back off very quickly.
%?? Talk about the same major points you made in the presentation here (i.e. TCP interactions) w/ pushback. Or in the evaluation...
UDP, however, behaves as expected---malicious hosts are expected to exhibit the same behaviour, given that congestion is their primary goal.

The performance characteristics of TCP under packet drop  are well-known and well-understood through the Mathis equation \cite{DBLP:journals/ccr/MathisSMO97}:
\begin{equation}
\operatorname{BW} = \frac{\operatorname{MSS}}{\operatorname{RTT}} \frac{C}{\sqrt{p}},
\end{equation}
for the \emph{maximum segment size} (MSS), \emph{round trip time} (RTT), a choice of constant $C \approx{} \sqrt{3/2}$ dependent on system assumptions, and the probability $p \in (0, 1]$ that a packet is lost.
%The lower bound on $p$ is deliberately open, as packet drop acts as a control signal for the TCP send window size and is an expected part of normal operation.
TCP flows in Mininet are found to obey this law, as expected.
This reveals that the original work \emph{does not capture this behaviour}, or else the initial experiment (UDP-like traffic when $n=2$) wouldn't have achieved similar results to those which were reported.

%\subsection{Synthesis}
%%It'd be really good if I could figure out a rough proof for the service property...
%%Okay... now say what this means for future viability! What changes can I make? Flow-specific metrics, of course! Increase granularity as much as I can!
%Both of these experiments suggest that the most viable way to enhance and deploy this MARL approach is to consider flows individually to maximise performance for legitimate hosts and to detect malicious hosts in a more principled manner.
%While this will add significant computational cost, this may be worked around with intelligent sampling and monitoring while taking multiple actions per-timestep \cite{DBLP:conf/hotnets/MaoAMK16}.
%
%?? Observations: training time lengthier by nature of emulated environment.
%?? Risk of going for a simulation (i.e. numeric)? Always going to be interesting behaviour that is missed out on (in theory), at the cost of training time/limits of simulation speed

@@

?? Other recent work in the prevention of DDoS? (i.e., non-RL).
\emph{Athena} \cite{DBLP:conf/dsn/LeeKSPY17} had it as oart of their evaluation, BGP-like rerouting techniques (specifically for transit-link attacks, but also for `normal' attacks) \cite{RoutingAroundCongestion}

@@

For categorising how RL fits into solving problems, I label works as direct- or indirect-control RL.
A \emph{direct-control} RL problem is one where the RL agent(s) are learning optimal control over a set of actions to act as the \emph{primary} defence or decision-maker---requiring measurements, reward functions and action sets tailored for this purpose.
?? Good Direct-control work? What this work is based on (which directly captures the spirit of ) \cite{DBLP:phd/ethos/Malialis14, DBLP:journals/eaai/MalialisK15}

An \emph{indirect-control} RL problem, where the role played by the agents is to act in service to \emph{another technique} responsible for decision-making, further optimising or generalising aspects of its operation beyond that of hand-coded heuristics.
%Indirect-control applications: link to all the juicy stuff in e.g. HotNets, data-driven networking, the works. \cite{DBLP:conf/hotnets/ValadarskySST17, DBLP:conf/hotnets/MaoAMK16}.
A past example includes learning when best to \emph{communicate} and share knowledge between \emph{hidden Markov model} anomaly detectors \cite{DBLP:conf/paisi/XuSH07}.
The position of this work is weakened by its reliance on the discredited `DARPA99' dataset \cite{DARPA-IDD, DBLP:conf/cisda/TavallaeeBLG09, DBLP:conf/sp/SommerP10}, but the idea itself is well-treated and this acts as a driver for improvements in this direction.
Outside of anomaly/intrusion detection, there has been growing interest in the use of reinforcement learning in data-driven networking, such as for intra-AS route optimisation \cite{DBLP:conf/hotnets/ValadarskySST17} and for resource-constrained process allocation \cite{DBLP:conf/hotnets/MaoAMK16}.
These works emphasise the necessity of ingenuity in effectively handling how states and actions are represented.

@@

It all ends here...

@@

?? Benefit of the more realistic emulation environ is that it is far closer in behaviour and architecture (i.e. viable) to a real SDN-enabled deployment, captures some dynamics which were otherwise hidden/lost by human ignorance. It also allows me to develop the system towards evolving traffic models where it is expected that RL should shine over and above standard ML techniques. THEN: Room to introduce/roll-in dynamic changepoint detection or adaptive exploration \cite{DBLP:conf/ki/Tokic10, DBLP:conf/ki/TokicP11, DBLP:conf/annpr/TokicP12}?

?? Overarching goal: works which \emph{respect the complexity of the network environment} (rather than na\"{i}ve simulation, blind ML applications etc.) and choose well-considered pathways to solution. \emph{Call-to-action}?

?? Security? I suspect that the very qualities that make inference difficult in IDS/IPS also increase the level of challenge an advanced threat must overcome.
System state which is dependent on many signals drawn from across a wide network, in concert with their burstiness and unpredictability, may have substantial effects on an attacker's capabilities.
?? Might want to mention it in related work above, but the recent attention on adversarial examples/tricking models needs to be looked into for RL. Poisoning attacks relevant for online techniques: old bounds exist \textcite{DBLP:journals/jmlr/KloftL10}, new stuff concerns collaborative learners \cite{DBLP:conf/acsac/ShenTS16}, nothing for rl. Hot topic in deep networks \cite{DBLP:conf/eurosp/PapernotMJFCS16, SpInMl}, but naturally still relevant with even linear approximations or exact tabular case due to limits of the PAC assumption.
?? evasion attacks by \textcite{DBLP:conf/sp/Carlini017}---all of these are computed by way of a general stochastic optimiser, such as \emph{Adam} \cite{DBLP:journals/corr/KingmaB14}. possible to apply something similar to our learned model to assess its security? would the suggested states even be valid? (i.e. since they're monotonically increasing for the most part).
